

\chapter{Skel Settings File}
Skel provides a collection of configurable options which are exposed in the
Skel settings file. The settings file is located in the user's home directory at
/.skel/settings.
The settings file consists of single line entries of the form
<name>=<value>.
Blank lines and lines starting with \# are ignored. Entries in the settings files are
case sensitive, and lower case is typically used for all names and yes/no values.
\section{Available Settings}
\begin{description}
  \item[deploy\_dir]
This is the directory into which the compiled applications will be
copied to be executed. This directory should be visible to the compute
nodes.
  \item[submit\_target]
This determines which submission script template will be used
by
{\tt skel\_submit}. Templates for titan and sith are included in this
release.
  \item[sleep\_before\_open (yes/no)]
Determines whether a short sleep statement is inserted in the generated 
code before the adios open call. Default is no.
  \item[barrier\_before\_open (yes/no)]
Determines whether an {\tt MPI\_Barrier} call is inserted into the 
generated code before {\tt adios\_open}. Default is yes.
  \item[barrier\_before\_access (yes/no)]
Determines whether an {\tt MPI\_Barrier} call is
inserted into the generated code before the sequence of
{\tt adios\_write} calls. Default is no.
  \item[barrier\_before\_close (yes/no)]
Determines whether an {\tt MPI\_Barrier}
call is inserted into the generated code before {\tt adios\_close}.
Default is no.
  \item[barrier\_after\_close (yes/no)]
Determines whether an {\tt MPI\_Barrier} call is inserted
into the generated code after {\tt adios\_close}. Default is no.
  \item[barrier\_after\_steps (yes/no)]
Determines whether a single {\tt MPI\_Barrier} call
is inserted into the generated code after the final
{\tt adios\_close}. Default is no.
  \item[use\_adios\_timing (yes/no)]
Determines whether a call is inserted to output
detailed timing information collected by ADIOS. For this to work, your
adios distribution must have been configured using
{\tt --enable-skel-timing}.
Default is no.
\end{description}

