%\vspace{30pt}
%\section*{{\Large \textbf{Introduction\label{HToc84890215}\label{HToc212016590}\label{HToc212016832}\label{HToc182553323}}}}
\section{Introduction}

%\vspace{24pt}
%\subsection*{{\large 1.1 }{\large \textbf{Goals}}}
\subsection{Goals}

%\vspace{10pt}
\leftskip=0pt
\parindent=0pt
As computational power has increased dramatically with the increase in the number 
of processors, input/output (IO) performance has become one of the most significant 
bottlenecks in today's high-performance computing (HPC) applications. With this 
in mind, ORNL and the Georgia Institute of Technology's Center for Experimental 
Research in Computer Systems have teamed together to design the Adaptive I/O System 
(ADIOS) as a componentization of the IO layer, which is scalable, portable, and 
efficient on different clusters or supercomputer platforms. We are also providing 
easy-to-use, high-level application program interfaces (APIs) so that application 
scientists can easily adapt the ADIOS library and produce science without diving 
too deeply into computer configuration and skills. \label{HToc84890216}\label{HToc212016591}\label{HToc212016833}\label{HToc182553324}

%\vspace{10pt}
%\subsection*{{\large 1.2 }{\large \textbf{What Is ADIOS?}}}
\subsection{What is ADIOS?}

%\vspace{10pt}
{\color{color01} ADIOS is a state-of-the-art componentization of the IO system 
that has demonstrated impressive IO performance results on leadership class machines 
and clusters; sometimes showing an improvement of more than 1000 times over well 
known parallel file formats. }ADIOS is essentially an I/O componentization of different 
I/O transport methods. This feature allows flexibility for application scientists 
to adopt the best I/O method for different computer infrastructures with very little 
modification of their scientific applications. ADIOS has a suite of simple, easy-to-use 
APIs. Instead of being provided as the arguments of APIs, all the required metadata 
are stored in an external Extensible Markup Language (XML) configuration file, 
which is readable, editable, and portable for most machines. \label{HToc212016592}\label{HToc212016834}\label{HToc84890217}\label{HToc182553325}

%\vspace{10pt}
%\subsection*{{\large 1.3 }{\large \textbf{The Basic ADIOS Group Concept}}}
\subsection{The Basic ADIOS Group Concept}

%\vspace{10pt}
\parindent=3pt
The ADIOS ``group'' is a concept in which input variables are tagged according 
to the functionality of their respective output files. For example, a common scientific 
application has checkpoint files prefixed with restart and monitoring files prefixed 
with diagnostics. In the XML configuration file, the user can define two separate 
groups with tag names of adios-group as ``restart'' and ``diagnostic.'' Each group 
contains a set of variables and attributes that need to be written into their respective 
output files. Each group can choose to have different I/O transport methods, which 
can be optimal for their I/O patterns.\label{HToc84890218}\label{HToc212016593}\label{HToc212016835}\label{HToc182553326}

%\vspace{10pt}
%\subsection*{{\large 1.4 }{\large \textbf{Other Interesting Features of ADIOS}}}
\subsection{Other Interesting Features of ADIOS}

%\vspace{10pt}
\parindent=0pt
ADIOS contains a new self-describing file format, BP. The BP file format was specifically 
designed to support delayed consistency, lightweight data characterization, and 
resilience. ADIOS also contains python scripts that allow users to easily write 
entire ``groups'' with the inclusion of one include statement inside their Fortran/C 
code. Another interesting feature of ADIOS is that it allows users to use multiple 
I/O methods for a single group. This is especially useful if users want to write 
data out to the file system, simultaneously capturing the metadata in a database 
method, and visualizing with a visualization method.

%\vspace{10pt}
The read API enables reading arbitrary subarrays of variables in a BP file and 
thus variables written out from N processor can be read in on arbitrary number 
of processors. ADIOS also takes care of the endianness problem at converting to 
the reader's architecture automatically at reading time. Matlab reader is included 
in the release while the VisIt parallel interactive visualization software can 
read BP files too (from version 2.0). 

%\vspace{10pt}
ADIOS is fully supported on Cray XT and IBM BlueGene/P computers as well as on 
Linux clusters and Mac OSX. \label{HToc84890219}\label{HToc212016594}\label{HToc212016836}\label{HToc182553327}

%\vspace{10pt}
%\subsection*{{\large 1.5 }{\large \textbf{Future ADIOS 2.0 Goals}}}
\subsection{Future ADIOS 2.0 Goals}

%\vspace{10pt}
One of the main goals for ADIOS 2.0 is to produce faster reads via indexing methods. 
Another goal is to provide more advanced data types via XML in ADIOS so that it 
will be compatible with F90/c/C++ structures/objects. 

%\vspace{10pt}
We will also work on the following advanced topics for ADIOS 2.0: 

\begin{itemize}
\item
A link to an external database for provenance recording.

\item
Autonomics through a feedback mechanism from the file system 
to optimize I/O performance. For instance, ADIOS can be adaptively changed from 
a synchronous to an asynchronous method or can decide when to write restart to 
improve I/O performance.

\item
A staging area for data querying, analysis, and in situ visualization.\label{HToc84890220}\label{HToc212016595}\label{HToc212016837}\label{HToc182553328}

\end{itemize}